Students' Alcohol Consumption impacts on academic Results on the example of a secondary Scool Maths Class in Portugal

### 2016-10-13 MSproject.Rmd
Adding descriptive statistics code

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```
Note that the `echo = FALSE` parameter was added to the code chunk to prevent printing of the R code that generated the plot.

@@

## Description of Dataset

```{r pressure, echo=FALSE}
sm <- read.csv("student-mat.csv", sep=';')
str(sm)
summary(sm[,"age"])
```

The dataset examined provides sample data of two secondary schools in Portugal, providing a total of 395 objects (students) and 33 variables (items on information on each student). A list of all variables and their description can be found in Appendix A. The data can be grouped into six areas of information:

#### 1. General Information
General information about the students include gender, address and age. From the descriptive statistics (section 3) we can derive that the age from the students in this dataset ranges from 15 to 22 years with an average of 16.7 years. The split between male and female students is .....and most student would live in an urban environment (.....), while % live in an rural environment.

@@

@@

#### 4. Social Life  
The social life section provieds information on activities the student is involved in outside of school. This includes

#### 5. Health

#### 6. Grades
The grades are split into three grades: the first period (G1), second period (G2) and final grade (G3). For the purpose of thy analysis, G1 and G2 will be neglected.

## Descriptive Statistics

You can include R code in the document as follows:

```{r cars}
summary(cars)
```

>>>>

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

@@

## Description of Dataset and Descriptive Statistics

```{r pressure, include=FALSE}
sm <- read.csv("student-mat.csv", sep=';')

# Descriptive statistics
library(ggplot2)

meanAge <- mean(sm[,'age']) #16.696
medianAge <- median(sm[,'age']) #17 
summary(sm[,'sex']) #F=208 #M=187
summary(sm[,'school']) #Gabriel Pereira = 349   #Mousinho da Silveira = 46
summary(sm[,'address']) #Rural = 77 #Urban = 307
summary(sm[,'famsize']) #Greater than 3 = 281 #Less than 3 = 114
summary(sm[,'Pstatus']) #A = 41 #T = 354
summary(sm[,'internet']) #no = 66 #yes=329
summary(sm[,'romantic']) #no=263 #yes=132

########################################################
# Blank Theme for Pie Charts below

blank_theme <- theme_minimal()+ theme(
    axis.title.x = element_blank(),
    axis.title.y = element_blank(),
    panel.border = element_blank(),
    panel.grid=element_blank(),
    axis.ticks = element_blank(),
    plot.title=element_text(size=14, face="bold")
  )
########################################################
# Sex
summary(sm[,'sex']) 
#F=208 #M=187

df_gender <- data.frame(Gender=c("Female", "Male"), value=c(208, 187))
pie_gender <- ggplot(data=df_sex, aes(x="", y=value, fill=Gender)) + 
  geom_bar(width = 1, stat = "identity") + 
  coord_polar("y", start=0) + 
  scale_fill_brewer() + blank_theme + 
  theme(axis.text.x=element_blank()) +
  geom_text(aes(y = value/3 + c(0, cumsum(value)[-length(value)]), label = (value)), size=4)

###
# Address
summary(sm[,'address']) 
#Rural = 77 #Urban = 307

df_address <- data.frame(Address=c("Rural", "Urban"), value=c(77, 307))
pie_address <- ggplot(data=df_address, aes(x="", y=value, fill=Address)) + 
  geom_bar(width = 1, stat = "identity") + 
  coord_polar("y", start=0) + 
  scale_fill_brewer() + blank_theme + 
  theme(axis.text.x=element_blank()) +
  geom_text(aes(y = value/3 + c(0, cumsum(value)[-length(value)]), label = (value)), size=4)

###
#Study Time

studyTime <- table(sm$studytime)
#  1   2   3   4 
# 105 198  65  27

df_studyTime <- data.frame(Study_Time=c('<2 hours', "2 to 5 hours", "5 to 10 hours", ">10 hours"), value=c(105, 198, 65, 27))
#weekly study time (numeric: 1 - <2 hours, 2 - 2 to 5 hours, 3 - 5 to 10 hours, or 4 - >10 hours)

bar_studytime <- ggplot(data=df_studyTime, aes(x="", y=value, fill=Study_Time)) + 
  geom_bar(width = 1, stat = "identity") + 
  scale_fill_brewer() + blank_theme + 
  theme(axis.text.x=element_blank()) +
  geom_text(aes(y = value/3 + c(0, cumsum(value)[-length(value)]), label = (value)), size=4)

###
# Failures
# number of past class failures (numeric: n if 1<=n<3, else 4)

failures <- table(sm$failures)
# 0   1   2   3 
# 312  50  17  16 
df_failures <- data.frame(Failures_Times=c("0", "1", "2", "3"), value=c(312, 50, 17, 16))

pie_failures <- ggplot(data=df_failures, aes(x="", y=value, fill=Failures_Times)) + 
  geom_bar(width = 1, stat = "identity") + 
  coord_polar("y", start=0) + 
  scale_fill_brewer() + blank_theme + 
  theme(axis.text.x=element_blank()) +
  geom_text(aes(y = value/3 + c(0, cumsum(value)[-length(value)]), label = (value)), size=4)

###
# Schools
summary(sm[,'school']) 
#Gabriel Pereira = 349   #Mousinho da Silveira = 46

df_school <- data.frame(Schools=c("Gabriel Pereira", "Mousinho da Silveira"), value=c(349,46))
bar_school <- ggplot(data=df_school, aes(x="", y=value, fill=Schools)) + 
  geom_bar(width = 1, stat = "identity") + 
  blank_theme + 
  theme(axis.text.x=element_blank()) +
  geom_text(aes(y = value/3 + c(0, cumsum(value)[-length(value)]), label = (value)), size=4)

###
# Father's Job
summary(sm[,'Fjob'])
ggplot(data=sm) + geom_bar(aes((sm[,'Fjob']))) + labs(x="Father's Job")
# at_home   health  other  services  teacher 
#   20       18      217      111       29

df_fjob <- data.frame(Group = c("Health", "Work at Home", "Teacher", "Servies", "Other"), value = c(18, 20, 29, 111, 217))
pie_fjob <- ggplot(data = df_fjob, aes(x="", y=value, fill=Group)) + 
  geom_bar(width = 1, stat = "identity") + 
  coord_polar("y", start=0) + 
  scale_fill_brewer() + blank_theme +
  theme(axis.text.x=element_blank()) +
  geom_text(aes(y = value/3 + c(0, cumsum(value)[-length(value)]), label = (value)), size=4) +
  ggtitle("Father's Job")

###
# Mother's Job
summary(sm[,'Mjob'])
ggplot(data=sm) + geom_bar(aes((sm[,'Mjob']))) + labs(x="Mother's Job")
# at_home   health  other  services  teacher 
# 59       34        141      103       58

df_mjob <- data.frame(Group = c("Health", "Teacher", "Work at Home", "Services", "Other"), value = c(34, 58, 59, 103, 141))
pie_mjob <- ggplot(data = df_mjob, aes(x="", y=value, fill=Group)) + 
  geom_bar(width = 1, stat = "identity") + 
  coord_polar("y", start=0) + 
  scale_fill_brewer() + blank_theme +
  theme(axis.text.x=element_blank()) +
  geom_text(aes(y = value/3 + c(0, cumsum(value)[-length(value)]), label = (value)), size=4) +
  ggtitle("Mother's Job")

###
# Student's Guardian
summary(sm[,'guardian'])
ggplot(data=sm) + geom_bar(aes(guardian)) + labs(x="Student's Guardian")
# father mother  other 
# 90    273     32

df_guardian <- data.frame(group = c("Father", "Mother", "Other"), value = c(90, 273, 32))
pie_guardian <- ggplot(data = df_guardian, aes(x="", y=value, fill=group)) + 
  geom_bar(width = 1, stat = "identity") + 
  coord_polar("y", start=0) + 
  scale_fill_brewer() + blank_theme +
  theme(axis.text.x=element_blank()) +
  geom_text(aes(y = value/3 + c(0, cumsum(value)[-length(value)]), label = (value)), size=4) +
  ggtitle("Guardian")

###############################################################################################

The dataset examined provides sample data of two secondary schools in Portugal, providing a total of 395 objects (students) and 33 variables (items on information on each student). A list of all variables and their description can be found in Appendix A. The data can be grouped into five areas of information:

#### 1. General Information
General information about the students include gender, address and age. From the descriptive statistics (section 3) we can derive that the age from the students in this dataset ranges from 15 to 22 years with an average of 16.7 years. Out of the total of 395, there are 52.7% female and 47.3% male students. Most students would live in an urban environment, while approx. 20% live in an rural environment.

```{r, echo=FALSE}
pie_gender

pie_address
```

@@

```{r, echo=FALSE}
pie_fjob

pie_mjob

pie_guardian
```

@@

```{r, echo=FALSE}
bar_school
bar_studytime

#### 4. Social Life and Health 
The social life section provieds information on activities the student is involved in outside of school. This includes

#### 5. Grades
The grades are split into three grades: the first period (G1), second period (G2) and final grade (G3). For the purpose of thy analysis, G1 and G2 will be neglected. 
```{r, echo=FALSE}
pie_failures
